---
title: "06-exercises-more-shiny"
author: "Charlie Reinhardt"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## 06 Exercises: More Shiny

Read the [UI Chapter of Mastering Shiny][id]

**Exercise 1.** In the tennis app, change the histogram input from a sliderInput() to a numericInput().

```{r}
library(tidyverse)
library(here)
```

```{r}
atp_df <- read_csv(here("data/atp_matches_2019.csv"))
wta_df <- read_csv(here("data/wta_matches_2019.csv"))
both_df <- bind_rows(atp_df, wta_df)

both_long <- both_df %>% pivot_longer(c(winner_name, loser_name))
```

```{r}
## only keep players who have player over 50 matches
both_n50 <- both_long %>% group_by(value) %>% count() %>%
  filter(n > 50)

## construct various statistics
major_tennis <- semi_join(both_long, both_n50, by = c("value"))
major_tennis <- major_tennis %>% mutate(w_svperc = 100 * w_1stIn / w_svpt,
                                        l_svperc = 100 * l_1stIn / l_svpt,
                                        w_firstwon = 100 * w_1stWon / w_1stIn,
                                        l_firstwon = 100 * l_1stWon / l_1stIn,
                                        w_secondwon = 100 * w_2ndWon / (w_svpt - w_1stIn),
                                        l_secondwon = 100 * l_2ndWon / (l_svpt - l_1stIn))

major_tennis_w <- major_tennis %>% filter(name == "winner_name")
major_tennis_l <- major_tennis %>% filter(name == "loser_name")

w_small <- major_tennis_w %>% select(value, winner_seed, w_ace, w_df, w_svperc,
                                     w_firstwon, w_secondwon) %>%
  rename(seed = winner_seed, ace = w_ace, df = w_df, svperc = w_svperc,
         firstwon = w_firstwon, secondwon = w_secondwon)

l_small <- major_tennis_l %>% select(value, loser_seed, l_ace, l_df, l_svperc, l_firstwon, l_secondwon)  %>%
  rename(seed = loser_seed, ace = l_ace, df = l_df, svperc = l_svperc,
         firstwon = l_firstwon, secondwon = l_secondwon)

df <- bind_rows(w_small, l_small) %>%
  rename(player = "value") 

df %>%
  filter(player == "Daniil Medvedev") %>%
  select(player, ace) %>%
  
  ggplot(., aes(x = ace)) + 
  geom_histogram()
```

```{r}
library(shiny)

ui <- fluidPage(
  selectizeInput(inputId = "playerchoice",
              label = "Choose a player", 
              choices = df$player),
  radioButtons(inputId = "varchoice",
                 label = "Choose a Statistic", 
                 choices = names(df)[c(3:7)]),
  numericInput("binnumber", label = "Choose a Number of Bins", 
               value = 10, min = 0, max = 50),
  plotOutput("hist")
)

server <- function(input, output, session) {
  # we need to wrap with reactive({}) if something depends on inputs
  # if it doesn't depend on input, we should just do it outside shinyapp
  new_df <- reactive({ 
    df %>%
      filter(player == input$playerchoice) %>%
      select(player, .data[[input$varchoice]])
  })
    
  output$hist <- renderPlot({
    ggplot(data = new_df(), aes(x = .data[[input$varchoice]])) + 
    geom_histogram(color = "black", fill = "white", bins = input$binnumber)
  })
    
}

shinyApp(ui, server)
```


**Exercise 2.** In the tennis app, change the histogram input back to sliderinput(). Then, apply 2.2.8 Exercise 3 in Mastering Shiny to the tennis app slider.
```{r}
# Couldn't figure out how to make the slider animate
library(shiny)

ui <- fluidPage(
  selectizeInput(inputId = "playerchoice",
              label = "Choose a player", 
              choices = df$player),
  radioButtons(inputId = "varchoice",
                 label = "Choose a Statistic", 
                 choices = names(df)[c(3:7)]),
  sliderInput("binnumber", label = "Choose a Number of Bins", 
                min = 0, max = 100, value = 15, step = 5),
  plotOutput("hist")
)

server <- function(input, output, session) {
  # we need to wrap with reactive({}) if something depends on inputs
  # if it doesn't depend on input, we should just do it outside shinyapp
  new_df <- reactive({ 
    df %>%
      filter(player == input$playerchoice) %>%
      select(player, .data[[input$varchoice]])
  })
    
  output$hist <- renderPlot({
    ggplot(data = new_df(), aes(x = .data[[input$varchoice]])) + 
    geom_histogram(color = "black", fill = "white", bins = input$binnumber)
  })
    
}

shinyApp(ui, server)
```

**Exercise 3.** In the SLU majors app, change the major input to radioButtons() so that the user can only select mathematics, statistics, or computer science as the major.
```{r}
library(tidyverse)
library(readxl)
library(here)
```

```{r}
df <- read_excel(here("data/slu_graduates_17_21.xlsx"))

## fixes error in the data
df <- df %>% mutate(across(everything(),
                           .fns = ~replace(., . ==  "STATS" , "STAT")))

df_long <- df %>% pivot_longer(3:8, names_to = "type", values_to = "discipline")
df_major <- df_long %>% 
  filter(type == "major1" | type == "major2" | type == "major3")

df_stat <- df_major %>% filter(discipline == "STAT") 
df_statfull <- semi_join(df_long, df_stat, by = "adm_id") %>%
  filter(type == "major1" |
           type == "major2" | 
           type == "major3")

df_nostat <- df_statfull %>% filter(discipline != "STAT" &
                                      !is.na(discipline)) %>%
  group_by(discipline) %>%
  summarise(nstudent = n()) %>%
  mutate(discipline = fct_reorder(discipline, nstudent))
ggplot(data = df_nostat, aes(x = discipline, y = nstudent)) +
  geom_col() +
  coord_flip()
```

```{r}
library(shiny)

df$major1

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      radioButtons(
        inputId = "majorchoice",  
        label = "Choose a Major", 
        choices = c("CS", "MATH", "STAT"))
    ),
    mainPanel(
      plotOutput(outputId = "majorplot")
    )
  )
)

server <- function(input, output, session) {
  
  df_update <- reactive({
    df_discipline <- 
      df_major %>% filter(discipline == input$majorchoice) 
    
    df_disciplinefull <-
      semi_join(df_long, df_discipline, by = "adm_id") %>%
      filter(type == "major1" |
               type == "major2" | 
               type == "major3")
    
    df_nodiscipline <- 
      df_disciplinefull %>% filter(discipline != input$majorchoice &
                                       !is.na(discipline)) %>%
      group_by(discipline) %>%
      summarise(nstudent = n()) %>%
      mutate(discipline = fct_reorder(discipline, nstudent))
  })
  
  
  output$majorplot <- renderPlot({
    ggplot(data = df_update(), aes(x = discipline, y = nstudent)) +
      geom_col() +
      coord_flip()
  })  
}

shinyApp(ui, server)
```


Exercise 4. In the SLU majors app, add a table output below the plot that shows the number of Female majors and the number of Male majors for a major that the user selects.
```{r}
library(tidyverse)
library(readxl)
library(here)
```

```{r}
df <- read_excel(here("data/slu_graduates_17_21.xlsx"))

## fixes error in the data
df <- df %>% mutate(across(everything(),
                           .fns = ~replace(., . ==  "STATS" , "STAT")))

df_long <- df %>% pivot_longer(3:8, names_to = "type", values_to = "discipline")
df_major <- df_long %>% 
  filter(type == "major1" | type == "major2" | type == "major3")

df_stat <- df_major %>% filter(discipline == "STAT") 
df_statfull <- semi_join(df_long, df_stat, by = "adm_id") %>%
  filter(type == "major1" |
           type == "major2" | 
           type == "major3")

df_nostat <- df_statfull %>% filter(discipline != "STAT" &
                                      !is.na(discipline)) %>%
  group_by(discipline) %>%
  summarise(nstudent = n()) %>%
  mutate(discipline = fct_reorder(discipline, nstudent))
ggplot(data = df_nostat, aes(x = discipline, y = nstudent)) +
  geom_col() +
  coord_flip()
```

```{r}
library(shiny)

df$major1

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      radioButtons(
        inputId = "majorchoice",  
        label = "Choose a Major", 
        choices = c("CS", "MATH", "STAT"))
    ),
    mainPanel(
      plotOutput(outputId = "majorplot"),
      tableOutput("table")
    )
  )
)

server <- function(input, output, session) {
  
  df_update <- reactive({
    df_discipline <- 
      df_major %>% filter(discipline == input$majorchoice) 
    
    df_disciplinefull <-
      semi_join(df_long, df_discipline, by = "adm_id") %>%
      filter(type == "major1" |
               type == "major2" | 
               type == "major3")
    
    df_summary <-
      df_disciplinefull %>% filter(discipline == "CS") %>%
      summarise(nfemale = sum(sex == "F"), nmale = sum(sex == "M"))
    
    df_nodiscipline <- 
      df_disciplinefull %>% filter(discipline != input$majorchoice &
                                       !is.na(discipline)) %>%
      group_by(discipline) %>%
      summarise(nstudent = n()) %>%
      mutate(discipline = fct_reorder(discipline, nstudent))
  })
  
  
  output$majorplot <- renderPlot({
    ggplot(data = df_update(), aes(x = discipline, y = nstudent)) +
      geom_col() +
      coord_flip()
  })
  
  output$table <- renderTable({
    df_summary
  })
}

shinyApp(ui, server)
```

```{r}
df_discipline <- 
  df_major %>% filter(discipline == "CS") 

df_disciplinefull <-
  semi_join(df_long, df_discipline, by = "adm_id") %>%
  filter(type == "major1" |
           type == "major2" | 
           type == "major3")

df_summary <-
  df_disciplinefull %>% filter(discipline == "CS") %>%
  summarise(nfemale = sum(sex == "F"), nmale = sum(sex == "M"))

df_nodiscipline <- 
  df_disciplinefull %>% filter(discipline != "CS" &
                                   !is.na(discipline)) %>%
  group_by(discipline) %>%
  summarise(nstudent = n()) %>%
  mutate(discipline = fct_reorder(discipline, nstudent))

df_summary
```


[id]: https://mastering-shiny.org/basic-ui.html "UI Chapter"