---
title: "05-tidyverse-review"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aside: Baseline of Zero?

```{r}
library(palmerpenguins)

ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point() + 
  xlim(c(0, 60)) + 
  ylim(c(0, 25))


```
When you're encoding length, baseline of zero is SUPER important. If you're encoding relative position (no length), baseline of zero is usually harmful.

## Section 5.1: Billboard
```{r}
library(billboard)
library(tidyverse)
head(wiki_hot_100s)
tail(wiki_hot_100s)
```

# Section 5.1.1 Exercises

Exercise 1
```{r}
wiki_hot_100s <- as_tibble(wiki_hot_100s)

wiki_top_15 <- 
wiki_hot_100s %>% 
  filter(year <= 2009 & year >= 2000) %>%
  group_by(artist) %>%
  summarise(nsongs = n()) %>%
  arrange(desc(nsongs)) %>%
  slice(1:15) %>%
  mutate(artist = fct_reorder(artist, nsongs))
  

ggplot(data = wiki_top_15, aes(x = artist, y = nsongs)) +
  geom_col() + 
  coord_flip()
```

Exercise 2. There is a minor flaw in the way that we counted up the number of hits for each artist. Examine the 2nd to last row of the original data set with tail() to look at this potential flaw. What do you find?
```{r}
tail(wiki_hot_100s)
```
Wiz Khalifa featuring Charlie Puth is identified as it's own artist, not as Wiz Khalifa and/or Charlie Puth

Exercise 3. Challenging. Fix the issue in Exercise 2. May want to skip this question and come back after completing the other exercises.
```{r}
wiki_hot_100s %>% 
  mutate(artist = str_remove(artist, pattern = " featuring .*")) %>% tail()
```

Exercise 4. Change the plot from Exercise 1 to be a Lollipop chart using this website as a reference. Why might the lollipop chart be better than a bar plot?
```{r}
wiki_top_15 %>%
  ggplot(data = ., mapping = aes(x = artist, y = nsongs)) + 
  geom_segment( aes(x = artist, xend=artist, y = 0, yend = nsongs)) + 
  geom_point() + 
  coord_flip()
```

To be honest, I don't quite understand. why a lollipop might be better
- Dr. Higham says higher data to ink ratio is good

Exercise 5. Use this website to customize the end points of your lollipop chart. If you have time, you can explore the other customization options. Make it look fancy!
```{r}
wiki_top_15 %>%
  ggplot(data = ., mapping = aes(x = artist, y = nsongs)) + 
  geom_segment( aes(x = artist, xend=artist, y = 0, yend = nsongs)) + 
  geom_point( size = 5, color = "red", shape = 18) + 
  coord_flip()
```

# Some webscraping
```{r}
library(rvest)
```

```{r}
get_wiki_100 <- function(year) {
  
  url <- paste0("https://en.wikipedia.org/wiki/Billboard_Year-End_Hot_100_singles_of_", year)
  
  h <- read_html(url)
  
  tab <- h %>% html_nodes("table")
  df <- tab[[1]] %>% html_table() %>% 
    mutate(year = year)
  
  return(df)
}
```

```{r}
get_wiki_100(year = 2017)
```

```{r}
library(purrr)
year_list <- list(2017, 2018, 2019, 2020, 2021)
df_all <- map(year_list, get_wiki_100)
```

```{r}
df_2017_present <- bind_rows(df_all)
```

```{r}
df_2017_present <- df_2017_present %>%
  mutate(Title = str_remove_all(Title, pattern = "\"")) %>% ## get rid of \ in title
  rename(no = No., 
         title = Title, 
         artist = `Artist(s)`) ## make column names match with billboard package

wiki_tibble <- as_tibble(wiki_hot_100s) %>% ## convert billboard data to tibble
  mutate(year = as.numeric(year),
         no = as.integer(no)) ## change variable types to match with scraped data
#> Warning in mask$eval_all_mutate(quo): NAs introduced by
#> coercion

hot100_df <- bind_rows(wiki_tibble, df_2017_present)
```

Exercise 6. Use the hot100_df to make either a bar plot or a lollipop chart of the most popular artists of the 2010s (2010 through 2019). It may be helpful to make this plot without looking back at the code for the 2000s plot until you get stuck.

```{r}
plot6 <-
hot100_df %>%
  filter(., year >= 2010, year <= 2019) %>%
  group_by(artist) %>%
  summarise(nsongs = n()) %>%
  arrange(desc(nsongs)) %>%
  mutate(artist = fct_reorder(artist, nsongs)) %>%
  slice(1:10) %>%
  
  ggplot(data=., aes(x = artist, y = nsongs)) + 
  geom_segment( aes(x = artist, xend=artist, y=0, yend=nsongs)) + 
  geom_point() + 
  coord_flip()
```

Exercise 7. Much of the code to scrape the data, using purrr to iterate over the scrape, and then combining the list of data frames to a single data frame may be new. It is not expected that you are able to write this code on your own, but you should have an overall understanding of what the code is doing. Write 2-3 sentences that summarizes the overall purpose of the rvest and purrr code.

Exercise 8. If you have time, use purrr and rvest to read different data from Wikipedia into R. For example, you might consider sports championship data, which will typically have a different Wikipedia page for each year or season.

## 5.2 More `tidyverse` Review: Happy Planet Index

```{r}
hpi_df <- read_csv("data/hpi-tidy.csv")
```

```{r}
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point()
```

```{r}
hpi_us <- hpi_df %>% filter(Country == "United States of America")
hpi_interest <- hpi_df %>% filter(Country == "United States of America" | 
                                    Country == "Sweden" |
                                    Country == "Russia")
```

```{r}
## install.packages("ggrepel")
library(ggrepel)
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point() +
  geom_label_repel(data = hpi_us, aes(label = Country)) +
  geom_point(data = hpi_us, size = 3, shape = 1) ## create a second point that is an open circle (shape = 1) with a larger size (size = 3) to surround the United States point on the scatterplot
```


# 5.2.2 `plotly` to Label Points Interactively
```{r}
## install.packages("plotly")
library(plotly)
```

```{r}
plot1 <- ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point()

ggplotly(plot1)
```

## messing with labels
```{r}
plot1 <- ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing, label = Country)) +
  geom_point()

ggplotly(plot1, tooltip = "label")
```

Exercise 2. Use the ggplotly() function on any other plot we’ve made so far.
```{r}
ggplotly(plot6)
```

## Exercise 3. 
What are some advantages of making a plot more interactive with `plotly`? What are some disadvantages?

Advantages: Interesting, allows user to explore data more
Disadvantages: Requires computer, adds extra "stuff", bad for making an argument about a graph

# 5.2.3 Themes and Color changes

```{r}
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point() +
  geom_label_repel(data = hpi_us, aes(label = Country)) +
  geom_point(data = hpi_us, size = 3, shape = 1) +
  labs(title = "Countries with a Higher Ecological Footprint Tend to Have Citizens with Higher Wellbeing", ## add title
    subtitle = "Wellbeing is on a 1-10 scale", ## add subtitle (smaller text size than the title)
    caption = "Data Source: http://happyplanetindex.org/countries", ## add caption to the bottom of the figure
    x = "Ecological Footprint", ## change x axis label
    y = "Wellbeing") ## change y axis label
```

# 5.2.4 Changing a color scale

## Exercise 4
In which of those 3 situation are we in for the World Region graph if we were to color by region?

unordered

## Exercise 5
```{r}
ggplot(data = hpi_df, aes(x = Footprint, y = HappyLifeYears, colour = Region)) +
  geom_point() +
  scale_colour_brewer(palette = "Set1")
```

## Exercise 6. 
Read the examples section of the Help file for ?scale_colour_viridis_d. What’s the difference between scale_colour_viridis_d(), ?scale_colour_viridis_c(), and scale_colour_viridis_b()?

`scale_colour_viridis_x` where `x = d, c, b` are for different types of data to color. Discrete (categorical groups), continuous (like a color gradient), and bins 

`_d` is for discrete values, `_c` is for continuous variables, and `_b` is for binning a continuous variable

# 5.2.5 Faceting

Faceting is more useful if:

* there are "more" categories and
* if there is a lot of overlap between groups




