---
title: "06-Mapping-Data"
author: "Charlie Reinhardt"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```



```{r}
## install.packages("maps")
library(maps)
library(tidyverse)
state_df <- ggplot2::map_data("state")
state_df %>% nrow()
```

```{r}
ggplot(data = state_df,
            mapping = aes(x = long, y = lat,
                          group = group)) +
  geom_polygon(fill = "white", color = "black") +
  coord_map() + 
  theme(axis.text = element_blank()) + # can use this for all the ticks, titles, etc
  theme_void() # or use this to get rid of all the extra stuff in a plot
  
```

```{r}
## install.packages("usdata")
library(usdata)
state_stats
state_stats <- state_stats %>% mutate(state = str_to_lower(state))
state_full <- left_join(state_df, state_stats, by = c("region" = "state"))

ggplot(data = state_full, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "black", aes(fill = coal)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) + # albers projection is a common projection
  theme_void() +
  scale_fill_viridis_b()
```

### Exercise 1

Which states had the fastest growth rate between 2000 and 2010? Make a variable for the percent change in population and then map this variable.

```{r}
state_full_2 <- 
  state_full %>% 
  mutate(pct_pop_change = ((pop2010 - pop2000) / pop2000) * 100)
```

```{r}
ggplot(data = state_full_2, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "black", aes(fill = pct_pop_change)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) + # albers projection is a common projection
  theme_void() +
  scale_fill_viridis_b()
```

### Exercise 2

To your map in Exercise 1, think about which type of colour scale makes the most sense to use (sequential, diverging, or unordered). Change the colour scale to match what makes the most sense.

```{r}
ggplot(data = state_full_2, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "black", aes(fill = pct_pop_change)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) + # albers projection is a common projection
  theme_void() +
  scale_fill_fermenter(direction = 1)
```


`scale_fill_brewer()` plots discrete variables/colors
`scale_fill_fermenter()` bins continuous variables into discrete colors
`scale_fill_distiller()` plots continuous colors on a color gradient


```{r}
active <- read_csv("https://raw.githubusercontent.com/iramler/stlawu_covid/main/slc_towns_active_cases.csv", n_max = 34)
tcases <- read_csv("https://raw.githubusercontent.com/iramler/stlawu_covid/main/slc_towns_total_cases.csv", n_max = 34)

active_long <- active %>% pivot_longer(5:ncol(active), names_to = "date",
                                       values_to = "cases")
## repeat for total cases
tcases_long <- tcases %>% pivot_longer(5:ncol(tcases), names_to = "date",
                                       values_to = "cases")

```

```{r}
library(lubridate)
covid_df <- left_join(tcases_long, active_long,
                      by = c("date", "Order", "NAME")) %>%
  mutate(date = mdy(date)) %>%
  rename(total_cases = cases.x,
         active_cases = cases.y) %>%
  mutate(total_cases = if_else(is.na(total_cases),
                               true = 0, false = total_cases),
         active_cases = if_else(is.na(active_cases),
                                      true = 0, false = active_cases))
  

covid_SLC <- covid_df %>% filter(NAME == "St. Lawrence County")
covid_sub <- covid_df %>% filter(NAME != "St. Lawrence County")
```

### Exercise 1

Make a line plot that shows the number of active_cases in all of St. Lawrence County over time.

```{r}
covid_SLC %>% filter(!is.na(Area.y) & !is.na(Population.y)) %>%
  mutate(month = month(date)) %>%
  
  ggplot(., aes(x = date, y = active_cases)) +
  geom_line()
```


```{r}
## install.packages("sf")
library(sf)
shp <- read_sf("data/SLC_Civil_Boundaries_SHP/slc.shp") %>% # basically read_csv
  st_transform(st_crs("+proj=longlat")) # specifies longitude and latitude coordinate system

ggplot(data = shp) +
  geom_sf() +
  theme_void()
```


```{r}
full_df <- left_join(shp, covid_sub, by = "NAME") %>%
  filter(date == max(date)) ## only plot cases on the most recent date
ggplot(data = full_df) +
  geom_sf(aes(fill = active_cases)) +
  theme_void()
```

### Exercise 2. 

Change the fill scale of the plot. Should you use an unordered, sequential, or diverging scale?

Sequential, because starts at 0 active cases and goes up from there
```{r}
full_df <- left_join(shp, covid_sub, by = "NAME") %>%
  filter(date == max(date)) ## only plot cases on the most recent date
ggplot(data = full_df) +
  geom_sf(aes(fill = active_cases)) +
  theme_void() + 
  scale_fill_distiller(type = "seq", palette = 7, direction = 1)
```

### Exercise 3. 

Change the colour scale so that active_cases are put into different bins with scale_fill_viridis_b(). What are some advantages and disadvantages of this?

```{r}
full_df <- left_join(shp, covid_sub, by = "NAME") %>%
  filter(date == max(date)) ## only plot cases on the most recent date
ggplot(data = full_df) +
  geom_sf(aes(fill = active_cases)) +
  theme_void() + 
scale_fill_viridis_b(n.breaks = 5) # can change number of bins in scale
```

Advantages: easier to differentiate between different active case numbers between towns, legend is easier on the eyes, simpler
Disadvantages: less specificity/minute differences

### Exercise 4. 

Explore the ?geom_sf_text() function and add the actual number of cases to the subregions in the plot, as is done on the SLC website.

```{r}
full_df <- left_join(shp, covid_sub, by = "NAME") %>%
  filter(date == max(date)) ## only plot cases on the most recent date
ggplot(data = full_df) +
  geom_sf(aes(fill = active_cases)) + 
  geom_sf_text(aes(label = active_cases), color = "blue") + # ugly but demonstrates the code
  theme_void() + 
  scale_fill_distiller(type = "seq", palette = 5, direction = 1)
```


## 6.2. Do we need a map?

### Exercise 1. 

Think of a second example where, even though the data is spatial, the best graph to show a particular point would not be a map.

In general, trends through time or relationships between two variables aren't best shown with a map

## 6.4 Leaflet

```{r}
## install.packages("devtools")
library(devtools)
## devtools::install_github("li-wen-li/uszipcodes")
library(uszipcodes)
```

```{r}
beers <- read_csv("data/breweries.csv") 
raw_zip <- uszipcodes::get_zip(beers$address)
beers$Zip <- as.integer(uszipcodes::clean_zip(raw_zip))

## only keep zip, lat, and long
zip_tab <- zip_table %>% dplyr::select(Zip, Latitude, Longitude)
beer_location <- inner_join(beers, zip_tab)
content <- beer_location %>%
  mutate(popup = paste0('<a href =', beer_location$website, '>',
                        beer_location$brewery_name, '</a>'))

```

```{r}
library(leaflet)

beer_map <- leaflet(beer_location) %>%
  setView(lng = -98.583, lat = 39.833, zoom = 4) %>% 
  addTiles() %>% 
  addProviderTiles(providers$Stamen.Watercolor) %>% 
  addMarkers(lng = beer_location$Longitude, lat = beer_location$Latitude,
             clusterOptions = markerClusterOptions(),
             popup = content$popup)
```

```{r}
beer_map
```

### Exercise 1. 

Why is `inner_join()` the most appropriate join function to use here in this example? What observations will an `inner_join()` get rid of from beers? from zip_tab?

`inner_join()` will remove (from beers) breweries without a valid, mappable zip code and will remove (from zip_tab) zipcodes without any breweries

### Exercise 2.

Examine this link (http://leaflet-extras.github.io/leaflet-providers/preview/) to look at various “provider” tiles. Choose one to change from Wikimedia and explain how the resulting map looks different.

It just changes the skin on the map. It can display different information depending on the context in which you want the graphic to be viewed (artistically, streets, geography, etc.)


