---
title: "05-quiz-inclass.Rmd"
author: "Charlie Reinhardt"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Quiz 5 In-Class

Complete this quiz in a .Rmd file. To turn in the quiz, push both a .Rmd file and a knitted .html file to your GitHub site.


“All work presented is my own. I have not communicated with or worked with anyone else on this quiz.”

Charlie Reinhardt

Collaboration Reminder: You may not communicate with or work with anyone else on this quiz, but you may use any of our course materials or materials on the Internet.



### Question 1 (7 points). 

Consider the following two bar plots using the palmerpenguins data set. The first is a plot of the penguin species while the second is a plot of the average bill length for each species.

```{r}
library(palmerpenguins)
library(tidyverse)
ggplot(data = penguins, aes(x = species)) +
  geom_bar() +
  labs(y = "Count")
```

```{r}
ggplot(data = penguins %>% group_by(species) %>%
         summarise(avg_length = mean(bill_length_mm, na.rm = TRUE)),
       aes(x = species, y = avg_length)) +
  geom_col() +
  labs(y = "Average Bill Length")
```

The first is an appropriate graph to construct, as it displays a count of different groups, which is what a bar chart does best. The second is not appropriate because it does not display the uncertainty/variation of the summary statistic in question and can mislead viewers of the graph.

### 
Question 2 (9 points).

Use the Happy Planet Index data set to construct a graph that does not properly show variability in the underlying data. Recall that some variables in this data set are `LifeExpectancy`, `Wellbeing`, `Footprint`, and `Region` of the world.

```{r}
library(here)
hpi_df <- read_csv(here("data/hpi-tidy.csv"))
```

```{r}
hpi_df %>% group_by(Region) %>%
  summarise(avg_life = mean(LifeExpectancy)) %>%
  mutate(Region = factor(Region)) %>%
  mutate(Region = fct_reorder(Region, avg_life)) %>%
  
  ggplot(., aes(x = Region, y = avg_life)) +
  geom_point(size = 2) + 
  geom_segment(aes(x = Region, xend = Region, y = 0, yend = avg_life)) + 
  coord_flip() + 
  labs(y = "Average Life Expectancy")
```


### Question 3 (7 points). 

Fix your graph from the previous question so that it does properly show variability in the underlying data.

```{r}
hpi_df %>% group_by(Region) %>%
  mutate(avg_life = mean(LifeExpectancy)) %>%
  ungroup() %>%
  mutate(Region = factor(Region)) %>%
  mutate(Region = fct_reorder(Region, avg_life)) %>%

  
  ggplot(., aes(x = Region, y = LifeExpectancy)) +
  geom_point(alpha = 0.5) +
  geom_point(aes(y = avg_life), size = 2, color = "red") +
  coord_flip()
  
```

