---
title: "06-quiz-inclass.Rmd"
author: "Charlie Reinhardt"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Complete this quiz in a .Rmd file. To turn in the quiz, push both a .Rmd file and a knitted .html file to your GitHub site.


“All work presented is my own. I have not communicated with or worked with anyone else on this quiz.”

Collaboration Reminder: You may not communicate with or work with anyone else on this quiz, but you may use any of our course materials or materials on the Internet.



Question 1 (6 points). Using the SLU Majors data set, give an example of a situation where you would construct a static graph instead of an interactive graph. Why would the static graph be better?

One example of a situation where you would prefer a static graph instead of an interactive graph could be when presenting an idea to a group of people. During the 'construction' of the Data Science major here at SLU, I have no doubt that the department gave a few presentations with some data and some graphs (potentially showing an increase in CS and stat majors at SLU, potentially showing major combination breakdowns) to administration, and presenting a static graph here would be FAR more beneficial than an interactive graph because it is easier to understand, easier to present, and does not hinge on the user finding the same conclusion that you did within an interactive graph.



Question 2 (17 points). Consider this version of the Tennis App from class:

```{r}
library(shiny); library(tidyverse)
atp_df <- read_csv("data/atp_matches_2019.csv")
wta_df <- read_csv("data/wta_matches_2019.csv")
both_df <- bind_rows(atp_df, wta_df)

both_long <- both_df %>% pivot_longer(c(winner_name, loser_name))
```

``` {r}
## only keep players who have player over 50 matches
both_n50 <- both_long %>% group_by(value) %>% count() %>%
  filter(n > 50)

## construct various statistics
major_tennis <- semi_join(both_long, both_n50, by = c("value"))
major_tennis <- major_tennis %>% mutate(w_svperc = 100 * w_1stIn / w_svpt,
                        l_svperc = 100 * l_1stIn / l_svpt,
                        w_firstwon = 100 * w_1stWon / w_1stIn,
                        l_firstwon = 100 * l_1stWon / l_1stIn,
                        w_secondwon = 100 * w_2ndWon / (w_svpt - w_1stIn),
                        l_secondwon = 100 * l_2ndWon / (l_svpt - l_1stIn))

major_tennis_w <- major_tennis %>% filter(name == "winner_name")
major_tennis_l <- major_tennis %>% filter(name == "loser_name")

w_small <- major_tennis_w %>% select(value, winner_seed, w_ace, w_df, w_svperc,
                                     w_firstwon, w_secondwon) %>%
  rename(seed = winner_seed, ace = w_ace, df = w_df, svperc = w_svperc,
         firstwon = w_firstwon, secondwon = w_secondwon)

l_small <- major_tennis_l %>% select(value, loser_seed, l_ace, l_df, l_svperc, l_firstwon, l_secondwon)  %>%
  rename(seed = loser_seed, ace = l_ace, df = l_df, svperc = l_svperc,
         firstwon = l_firstwon, secondwon = l_secondwon)

df <- bind_rows(w_small, l_small) %>%
  rename(player = "value")
df

var_choices <- names(df)[3:7]
```

```{r}
library(shiny)

ui <- fluidPage(
  sidebarLayout(sidebarPanel(
    selectizeInput("playerchoice",
                   label = "Choose a Player", choices = levels(factor(df$player)),
                   selected = "Aryna Sabalenka"),
    radioButtons("varchoice", label = "Choose a Statistic",
                 choices = var_choices),
    selectizeInput("playerchoice2", 
                   label = "Choose another Player",
                   choices = levels(factor(df$player)))
    ),
   mainPanel(plotOutput("histgraph"),
             plotOutput("boxgraph"))
  )
)

server <- function(input, output, session) {
  
  df_sub <- reactive({
    df %>% group_by(player) %>% 
      filter(player == input$playerchoice || player == input$playerchoice2)
  })
  
  
  hist_plot <- reactive({
    # ggplot(df_sub(), aes_string(x = input$varchoice)) +
    # geom_histogram(colour = "black", fill = "white", bins = 15)
    ggplot(df_sub(), aes(x = .data[[input$varchoice]])) +
      geom_histogram(colour = "black", fill = "white", bins = 15)
  })
  
  output$histgraph <- renderPlot({
    hist_plot()
  })
  
  box_plot <- reactive({
      ggplot(data = df_sub(), aes(x = player, y = .data[[input$varchoice]])) + 
      geom_boxplot()
  })
  
  output$boxgraph <- renderPlot({
    box_plot()
  })
}

shinyApp(ui, server)
```

```{r}
## workspace
df %>% 
  group_by(player) %>%
  filter(player == "Daniil Medvedev" || player == "Jeremy Chardy") %>%
  
  
  ggplot(data = ., aes(x = player, y = ace)) + 
  geom_boxplot()
```

